Comparison of Fast, Hybrid Imaging Architectures for Multi-scale, Hierarchical Aperture Arrays

### 2024-10-18 ms.tex
Update sections 5.1 and 5.2

\textbf{Let $\widetilde{E}_{a_m}^{p}$ represent the calibrated and potentially noise-weighted electric field in polarisation state, $p$, measured by a station element, $a_m$, in station, $m$. Then, such electric fields from individual station elements}
% The calibrated electric fields, $\widetilde{E}_{a_m}^p$, recorded by individual elements, $a_m$, within a station, $m$, and polarisation state, $p$,
can be phase-coherently superposed \textbf{to obtain the electric field in polarisation state, $\alpha$,} towards any desired direction as
\begin{align}
    \widetilde{\mathcal{E}}_m^\alpha(\hat{\boldsymbol{s}}_k) &= \sum_{a_m} \sum_p  \widetilde{\mathcal{W}}_{a_m}^{{\alpha p}^*}(\hat{\boldsymbol{s}}_k) \, \widetilde{E}_{a_m}^p \, e^{i\frac{2\pi}{\lambda} \hat{\boldsymbol{s}}_k\cdot\boldsymbol{r}_{a_m}} \, , \label{eqn:intra-station-pol-hol-img-expl}
\end{align}
where, $\hat{\boldsymbol{s}}_k$ denotes the direction of the superposed beam, $k$,
% $\widetilde{E}_{a_m}^{p}$ represents the calibrated and potentially noise-weighted electric field measured by the station element, $a_m$,
and $\widetilde{\mathcal{W}}_{a_m}^{{\alpha p}^*}(\hat{\boldsymbol{s}}_k)$ denotes a complex-valued directional weighting\footnote{One can choose $\widetilde{\mathcal{W}}_{a_m}^{\alpha p}(\hat{\boldsymbol{s}}_k)=\mathcal{W}_{a_m}^{p\alpha}(\hat{\boldsymbol{s}}_k)$, the directional electric field sensitivity, if the signal-to-noise ratio is to be optimised. But it can also be generically chosen depending on the property desired of the estimator \cite[][]{Morales2011}.} \textbf{with the superscript indices $\alpha, p$, representing the contribution of polarisation state, $p$ in the station element, $a_m$, to polarisation state, $\alpha$, towards direction, $\hat{\boldsymbol{s}}_k$. The * denotes complex conjugate.}

The polarised intensity in the \textbf{beamformed} pixel is then obtained by
\begin{align}
    \widetilde{\mathcal{I}}^{\alpha\beta}_m(\hat{\boldsymbol{s}}_k) &= \left\langle \widetilde{\mathcal{E}}_m^\alpha(\hat{\boldsymbol{s}}_k) \,  \widetilde{\mathcal{E}}_m^{\beta^*}(\hat{\boldsymbol{s}}_k) \right\rangle \, , \label{eqn:intra-station-opt-pol-img-outprod}
\end{align}
where, angular brackets denote a temporal averaging across an interval of $t_\textrm{acc}$. \textbf{Superscripts $\alpha \beta$ in $\widetilde{\mathcal{I}}^{\alpha\beta}_m(\hat{\boldsymbol{s}}_k)$ index the four pairwise combinations of polarisation states of the intensity.} Even though it is an outer product over the polarisation axis, it will be referred to as a ``squaring'' operation, hereafter, for convenience because of what it reduces to if only a single polarisation is measured.

@@

Figure~\ref{fig:1D-incoherent-compcost-BF-LAMBDA-I} shows a breakdown of the computational cost per station per voxel for voltage beamforming as a function of the different station parameters. In each panel, all parameters except the one on the $x$-axis are kept fixed at the characteristic values of the \texttt{LAMBDA-I}, \textbf{\texttt{SKA-low}, and \texttt{SKA-low-core}} stations as highlighted in cyan. The total cost is dominated by the beamforming in Equation~(\ref{eqn:intra-station-pol-hol-img-expl}) denoted by black dashed lines. Beamforming cost per pixel scales linearly with the number of elements per station. The costs of squaring and temporal averaging in Equation~(\ref{eqn:intra-station-opt-pol-img-outprod}) are subdominant. Because they are performed on a per-pixel basis and not dependent on $N_\textrm{eps}$, they remain constant across $N_\textrm{eps}$. And because the number of independent pixels covering the field of view scales as $n_\textrm{bs} \simeq (D_\textrm{s}/D_\textrm{e})^2$, the cost per independent pixel remains constant across $D_\textrm{s}$ and $D_\textrm{e}$. All these operations have to be performed at every time step, $\delta t$, and hence remain constant regardless of $t_\textrm{acc}$. Although spatial averaging across multiple stations (orange dot-dashed line) is not a station-level operation, it is shown to emphasise that it constitutes only a negligible cost relative to the other components, noting that it needs to be only performed on a slower cadence of $t_\textrm{acc}$ on a per-pixel basis, and thus scales inversely with $t_\textrm{acc}$.

@@

Figure~\ref{fig:multidim-incoherent-compcost-BF-LAMBDA-I} shows a covariant view of the total computational cost density per station per voxel for the beamforming architecture taken two station parameters at a time while keeping the rest fixed at the nominal values of a \texttt{LAMBDA-I}, \textbf{\texttt{SKA-low}, and a \texttt{SKA-low-core}} station (grey dashed lines). The variation in the cost per voxel is dominated by the beamforming cost and exhibits the same trends observed in Figure~\ref{fig:1D-incoherent-compcost-BF-LAMBDA-I}, that is, linear in $N_\textrm{eps}$ while remaining constant with $D_\textrm{e}$ and $D_\textrm{s}$.

@@

Forming independent beams that simultaneously fill the entire field of view can be achieved by a two-dimensional spatial Fast Fourier Transform (FFT) of the measurements on the aperture plane. This can be enabled even for an irregular layout of station elements by gridding the calibrated electric fields measured by the station elements. The Optimal Map Making \citep[OMM;][]{Tegmark1997a} formalism adopted by \citet{Morales2011} is the basis of the architecture called E-field Parallel Imaging Correlator \cite[EPIC;][]{Thyagarajan+2017},
\begin{align}
    \widetilde{\mathcal{E}}_m^\alpha(\hat{\boldsymbol{s}}_k) &= \sum_j \delta^2 \boldsymbol{r}_j \, e^{i\frac{2\pi}{\lambda} \hat{\boldsymbol{s}}_k\cdot\boldsymbol{r}_j} \left(\sum_{a_m} \sum_p \widetilde{W}_{a_m}^{\alpha p*}(\boldsymbol{r}_j-\boldsymbol{r}_{a_m}) \, \widetilde{E}_{a_m}^p \right) \, . \label{eqn:intra-station-pol-hol-img-epic}
\end{align}
The expression within the parenthesis represents the gridding operation, and the outer summation denotes the Fourier transform implemented via FFT.

@@

Application of the FFT will have the effect of simultaneously beamforming over the entire field of view, $\Omega_\textrm{e}$. The convolution with $\widetilde{W}_{a_m}^{\alpha p*}(\boldsymbol{r})$ in the aperture, which is the Fourier transform equivalent of $\widetilde{\mathcal{W}}_{a_m}^{{\alpha p}^*}(\hat{\boldsymbol{s}}_k)$, has several purposes. Firstly, it can be chosen to optimise specific properties in the synthesised image. For example, choosing $\widetilde{W}_{a_m}^{\alpha p*}(\boldsymbol{r})=W_{a_m}^{\alpha p*}(\boldsymbol{r})$, \textbf{the complex conjugate of} the holographic illumination pattern of the antenna element, $a_m$, will optimise the signal-to-noise ratio in the image \citep{Morales2011}. \textbf{Secondly, it can incorporate $w$-projection for non-coplanar element spacings \citep{Cornwell+2008}, direction-dependent ionospheric effects, wide-field distortions from refractive and scintillating atmospheric distortions, etc. \citep{Morales+2009,Morales2011}.} Finally, it also acts as gridding convolution transforming data from discrete and arbitrary locations onto a regular grid, thereby facilitating the application of a two-dimensional spatial FFT. Finally, the polarised intensities are obtained by application of Equation~(\ref{eqn:intra-station-opt-pol-img-outprod}).

@@

Figure~\ref{fig:1D-incoherent-compcost-EPIC-LAMBDA-I} shows a breakdown of the costs per station per voxel of the EPIC architecture, with nominal values of a \texttt{LAMBDA-I} station marked in cyan. The dominant cost is from the two-dimensional FFT (black dashed lines) in Equation~\ref{eqn:intra-station-pol-hol-img-epic}.
% or equivalently, Equation~\ref{eqn:intra-station-pol-hol-img-epic-LA}.
Because the spatial FFT cost depends only on the grid dimensions, it is independent of $N_\textrm{eps}$ as the grid can hold \textbf{as many} $N_\textrm{eps}$ elements as physically possible. This spatial FFT scales roughly as $(\gamma_\textrm{s} \, D_\textrm{s}/D_\textrm{e})^2\log_2(\gamma_\textrm{s}\, D_\textrm{s}/D_\textrm{e})^2$, where, $\gamma_\textrm{s}$ is a constant padding factor. \textbf{$\gamma_\textrm{s}$ is used to control the pixel scale in the image. For example, $\gamma_\textrm{s}=2$ will provide identical pixel scale and image size as that formed from gridded visibilities.} Notably, because the spatial FFT produces an image over the entire field of view at every independent pixel instantly, it is not a pixelwise operation, and therefore, its cost per pixel is $\sim \log_2(D_\textrm{s}/D_\textrm{e})$, showing a weak dependence on $D_\textrm{s}$ and $D_\textrm{e}$. The pixelwise squaring (pink dot-dashed lines) and temporal averaging (blue dot-dashed lines) costs in Equation~\ref{eqn:intra-station-opt-pol-img-outprod}
% {eqn:intra-station-opt-polarimetric-direct-img-LA}
are subdominant. These being per pixel operations, the cost per voxel remains constant with $N_\textrm{eps}$, $D_\textrm{s}$ and $D_\textrm{e}$ similar to the voltage beamforming architecture. The gridding (black dotted lines) with weights in Equation~(\ref{eqn:intra-station-pol-hol-img-epic})
% and (\ref{eqn:intra-station-pol-hol-img-epic-LA})
is a per-element operation and depends only on $N_\textrm{eps}$ \textbf{and the size of the gridding kernel, $N_\textrm{ke}$. Here, $N_\textrm{ke}=1$ is chosen to provide unaliased full field-of-view images. The incorporation of $w$-projection in the gridding kernel will correspondingly increase $N_\textrm{ke}$. Nevertheless, the gridding cost}, scales linearly with $N_\textrm{eps}$ and inverse squared with $D_\textrm{s}$ per voxel. Overall, the gridding cost is insignificant. Again, the spatial averaging of pixels across multiple stations (orange dot-dashed line) is not a station-level operation, but is only shown to emphasise that it operates on a much slower cadence of $t_\textrm{acc}$, scaling inversely with $t_\textrm{acc}$, and adds only negligible cost to the overall cost budget.

@@

Figure~\ref{fig:multidim-incoherent-compcost-EPIC-LAMBDA-I} shows a covariant view of the total computational cost density per station per voxel for the EPIC architecture taken two station parameters at a time while keeping the rest fixed at the nominal values of a \texttt{LAMBDA-I} station denoted by grey dashed lines. The variation in the cost per voxel is dominated by the spatial FFT cost which is relatively constant in all these parameters and displays the same trends observed in Figure~\ref{fig:1D-incoherent-compcost-EPIC-LAMBDA-I}, namely, nearly constant in $N_\textrm{eps}$, $D_\textrm{e}$ and $D_\textrm{s}$. Figure~\ref{fig:multidim-incoherent-relative-compcost-EPIC-LAMBDA-I} shows the ratio of the computational cost of the EPIC architecture relative to the voltage beamforming architecture in Figure~\ref{fig:multidim-incoherent-compcost-BF-LAMBDA-I}. It is clearly noted that the cost of the EPIC architecture is significantly less than that of voltage beamforming for $N_\textrm{eps} \gtrsim 100$, and continues to decrease with increasing $N_\textrm{eps}$, implying that for full field of view imaging, EPIC has a clear and significant computational advantage over voltage beamforming as $N_\textrm{eps}$ and packing density of elements in a station increase.

@@

\textbf{The EPIC architecture will be relevant for large cosmological arrays like PUMA that are considering a FFT-based correlator architecture. Obtaining visibilities from an EPIC architecture will involve an additional step of an inverse-FFT or an inverse-DFT back to the aperture plane from the accumulated images (not included in Figure~\ref{fig:1D-incoherent-compcost-EPIC-LAMBDA-I}). However, the inverse-FFT or DFT operation needs to be performed only once every accumulation interval, and is thus not expected to add significantly to the total cost budget.}

>>>>

\textbf{Let $\widetilde{E}_{a_m}^{p}$ represent the calibrated and potentially noise-weighted electric field in polarisation state, $p$, measured by a station element, $a_m$, in station, $m$. Then, such electric fields from individual station elements}
% The calibrated electric fields, $\widetilde{E}_{a_m}^p$, recorded by individual elements, $a_m$, within a station, $m$, and polarisation state, $p$,
can be phase-coherently superposed \textbf{to obtain the electric field in polarisation state, $\alpha$,} towards any desired direction as
\begin{align}
    \widetilde{\mathcal{E}}_m^\alpha(\hat{\boldsymbol{s}}_k) &= \sum_{a_m} \sum_p  \widetilde{\mathcal{W}}_{a_m}^{{\alpha p}^*}(\hat{\boldsymbol{s}}_k) \, \widetilde{E}_{a_m}^p \, e^{i\frac{2\pi}{\lambda} \hat{\boldsymbol{s}}_k\cdot\boldsymbol{r}_{a_m}} \, , \label{eqn:intra-station-pol-hol-img-expl}
\end{align}
where, $\hat{\boldsymbol{s}}_k$ denotes the direction of the superposed beam, $k$,
% $\widetilde{E}_{a_m}^{p}$ represents the calibrated and potentially noise-weighted electric field measured by the station element, $a_m$,
and $\widetilde{\mathcal{W}}_{a_m}^{{\alpha p}^*}(\hat{\boldsymbol{s}}_k)$ denotes a complex-valued directional weighting\footnote{One can choose $\widetilde{\mathcal{W}}_{a_m}^{\alpha p}(\hat{\boldsymbol{s}}_k)=\mathcal{W}_{a_m}^{p\alpha}(\hat{\boldsymbol{s}}_k)$, the directional electric field sensitivity, if the signal-to-noise ratio is to be optimised. But it can also be generically chosen depending on the property desired of the estimator \cite[][]{Morales2011}.} \textbf{with the superscript indices $\alpha, p$, representing the contribution of polarisation state, $p$ in the station element, $a_m$, to polarisation state, $\alpha$, towards direction, $\hat{\boldsymbol{s}}_k$. The * denotes complex conjugate. Equation~(\ref{eqn:intra-station-pol-hol-img-expl}) resembles a Discrete Fourier Transform (DFT) of the calibrated Electric field measurement after applying a complex weighting.}

The polarised intensity in the \textbf{beamformed} pixel is then obtained by
\begin{align}
    \widetilde{\mathcal{I}}^{\alpha\beta}_m(\hat{\boldsymbol{s}}_k) &= \left\langle \widetilde{\mathcal{E}}_m^\alpha(\hat{\boldsymbol{s}}_k) \,  \widetilde{\mathcal{E}}_m^{\beta^*}(\hat{\boldsymbol{s}}_k) \right\rangle \, , \label{eqn:intra-station-opt-pol-img-outprod}
\end{align}
where, angular brackets denote a temporal averaging across an interval of $t_\textrm{acc}$. \textbf{Superscripts $\alpha \beta$ in $\widetilde{\mathcal{I}}^{\alpha\beta}_m(\hat{\boldsymbol{s}}_k)$ result from the outer product of indices $\alpha$ and $\beta$, and thus index the four pairwise combinations of polarisation states of the intensity.} Even though it is an outer product over the polarisation axis, it will be referred to as a ``squaring'' operation, hereafter, for convenience because of what it reduces to if only a single polarisation is measured.

@@

Figure~\ref{fig:1D-incoherent-compcost-BF-LAMBDA-I} shows a breakdown of the computational cost per station per voxel for voltage beamforming as a function of the different station parameters. In each panel, all parameters except the one on the $x$-axis are kept fixed at the characteristic values of the \texttt{LAMBDA-I}, \textbf{\texttt{SKA-low}, and \texttt{SKA-low-core}} stations as highlighted in cyan. The total cost is dominated by the \textbf{DFT} beamforming in Equation~(\ref{eqn:intra-station-pol-hol-img-expl}) denoted by black dashed lines. \textbf{The DFT} beamforming cost per pixel scales linearly with the number of elements per station. The costs of squaring and temporal averaging in Equation~(\ref{eqn:intra-station-opt-pol-img-outprod}) are subdominant. Because the \textbf{squaring and temporal averaging} are performed on a per-pixel basis and not dependent on $N_\textrm{eps}$, they remain constant across $N_\textrm{eps}$. And because the number of independent pixels covering the field of view scales as $n_\textrm{bs} \simeq (D_\textrm{s}/D_\textrm{e})^2$, the cost per independent pixel remains constant across $D_\textrm{s}$ and $D_\textrm{e}$. All these operations have to be performed at every time step, $\delta t$, and hence remain constant regardless of $t_\textrm{acc}$. Although spatial averaging across multiple stations (orange dot-dashed line) is not a station-level operation, it is shown to emphasise that it constitutes only a negligible cost relative to the other components, noting that it needs to be only performed on a slower cadence of $t_\textrm{acc}$ on a per-pixel basis, and thus scales inversely with $t_\textrm{acc}$.

@@

Figure~\ref{fig:multidim-incoherent-compcost-BF-LAMBDA-I} shows a covariant view of the total computational cost density per station per voxel \textbf{(after summing the DFT, squaring, and averaging components)} for the beamforming architecture taken two station parameters at a time while keeping the rest fixed at the nominal values of a \texttt{LAMBDA-I}, \textbf{\texttt{SKA-low}, and a \texttt{SKA-low-core}} station (grey dashed lines). The variation in the cost per voxel is dominated by the beamforming cost and exhibits the same trends observed in Figure~\ref{fig:1D-incoherent-compcost-BF-LAMBDA-I}, that is, linear in $N_\textrm{eps}$ while remaining constant with $D_\textrm{e}$ and $D_\textrm{s}$.

@@

\textbf{Instead of DFT beamforming at discrete arbitrary locations as described above,} forming independent beams that simultaneously fill the entire field of view can be achieved by a two-dimensional spatial Fast Fourier Transform (FFT) of the \textbf{calibrated electric field} measurements on the aperture plane. \textbf{Typically, this has been associated with a regularly gridded array layout \citep{Daishido+1991,Otobe+1994,Tegmark+2009,Tegmark+2010,Foster+2014,Masui+2019}. However, it} can be enabled even for an irregular layout of station elements by gridding the calibrated electric fields measured by the station elements. The Optimal Map Making \citep[OMM;][]{Tegmark1997a} formalism adopted by \citet{Morales2011} is the basis of the architecture called E-field Parallel Imaging Correlator \cite[EPIC;][]{Thyagarajan+2017},
\begin{align}
    \widetilde{\mathcal{E}}_m^\alpha(\hat{\boldsymbol{s}}_k) &= \sum_j \delta^2 \boldsymbol{r}_j \, e^{i\frac{2\pi}{\lambda} \hat{\boldsymbol{s}}_k\cdot\boldsymbol{r}_j} \left(\sum_{a_m} \sum_p \widetilde{W}_{a_m}^{\alpha p*}(\boldsymbol{r}_j-\boldsymbol{r}_{a_m}) \, \widetilde{E}_{a_m}^p \right) \, . \label{eqn:intra-station-pol-hol-img-epic}
\end{align}
The expression within the parenthesis represents the gridding operation \textbf{of each calibrated electric field measurement at an arbitrary location, $\boldsymbol{r}_{a_m}$, of element $a_m$ onto a common grid at locations, $\boldsymbol{r}_j$, using a gridding kernel, $\widetilde{W}_{a_m}^{\alpha p*}(\boldsymbol{r}_j)$ corresponding to that element.} The outer summation denotes the Fourier transform \textbf{of the gridded electric fields} implemented via FFT.

@@

Application of the FFT will have the effect of simultaneously beamforming over the entire field of view, $\Omega_\textrm{e}$. The convolution with $\widetilde{W}_{a_m}^{\alpha p*}(\boldsymbol{r})$ in the aperture, which is the Fourier transform equivalent of $\widetilde{\mathcal{W}}_{a_m}^{{\alpha p}^*}(\hat{\boldsymbol{s}}_k)$, has several purposes. Firstly, it can be chosen to optimise specific properties in the synthesised image. For example, choosing $\widetilde{W}_{a_m}^{\alpha p*}(\boldsymbol{r})=W_{a_m}^{\alpha p*}(\boldsymbol{r})$, \textbf{the complex conjugate of} the holographic illumination pattern of the antenna element, $a_m$, will optimise the signal-to-noise ratio in the image \citep{Morales2011}. \textbf{Secondly, it can incorporate $w$-projection for non-coplanar element spacings \citep{Cornwell+2008}, direction-dependent ionospheric effects, wide-field distortions from refractive and scintillating atmospheric distortions, etc. \citep{Morales+2009,Morales2011}.} Finally, it also acts as gridding convolution transforming data from discrete and arbitrary locations onto a regular grid, thereby facilitating the application of a two-dimensional spatial FFT. Finally, the polarised intensities are obtained by application of Equation~(\ref{eqn:intra-station-opt-pol-img-outprod}), \textbf{that is pixelwise ``squaring'' (outer product of polarisation states) and temporal averaging.}

@@

Figure~\ref{fig:1D-incoherent-compcost-EPIC-LAMBDA-I} shows a breakdown of the costs per station per voxel of the EPIC architecture, with nominal values of a \texttt{LAMBDA-I}, \textbf{\texttt{SKA-low}, or a \texttt{SKA-low-core}} station marked in cyan. The dominant cost is from the two-dimensional FFT (black dashed lines) in Equation~\ref{eqn:intra-station-pol-hol-img-epic}.
% or equivalently, Equation~\ref{eqn:intra-station-pol-hol-img-epic-LA}.
The spatial FFT cost depends only on the grid dimensions. So, it is independent of $N_\textrm{eps}$ as the grid can hold \textbf{as many} $N_\textrm{eps}$ elements as physically possible. The \textbf{computational cost of the} spatial FFT scales roughly as $(\gamma_\textrm{s} \, D_\textrm{s}/D_\textrm{e})^2\log_2(\gamma_\textrm{s}\, D_\textrm{s}/D_\textrm{e})^2$, where, $\gamma_\textrm{s}$ is a constant padding factor. \textbf{$\gamma_\textrm{s}$ is used to control the pixel scale in the image. For example, $\gamma_\textrm{s}=2$ (adopted in this paper) will provide identical pixel scale and image size as that formed from gridded visibilities.} Notably, because the spatial FFT produces an image over the entire field of view at every independent pixel instantly, it is not a pixelwise operation, and therefore, its cost per pixel scales as $\sim \log_2(D_\textrm{s}/D_\textrm{e})$, which exhibits a weak dependence on $D_\textrm{s}$ and $D_\textrm{e}$. The squaring (pink dot-dashed lines) and temporal averaging (blue dot-dashed lines) \textbf{in Equation~\ref{eqn:intra-station-opt-pol-img-outprod} operate on a per-pixel basis. Thus, those} costs per voxel remain constant with $N_\textrm{eps}$, $D_\textrm{s}$ and $D_\textrm{e}$ similar to the voltage beamforming architecture. And they 
% {eqn:intra-station-opt-polarimetric-direct-img-LA}
are subdominant. The gridding (black dotted lines) with weights in Equation~(\ref{eqn:intra-station-pol-hol-img-epic})
% and (\ref{eqn:intra-station-pol-hol-img-epic-LA})
is a per-element operation and depends only on $N_\textrm{eps}$ \textbf{and the size of the gridding kernel, $N_\textrm{ke}$. Here, $N_\textrm{ke}=1$ is chosen to provide unaliased full field-of-view images. The incorporation of $w$-projection in the gridding kernel will correspondingly increase $N_\textrm{ke}$. Nevertheless, the gridding cost}, scales linearly with $N_\textrm{eps}$ and inverse squared with $D_\textrm{s}$ per voxel. Overall, the gridding cost is insignificant. Again, the spatial averaging of pixels across multiple stations (orange dot-dashed line) is not a station-level operation, but is only shown to emphasise that it operates on a much slower cadence of $t_\textrm{acc}$, scaling inversely with $t_\textrm{acc}$, and adds only negligible cost to the overall cost budget.

@@

Figure~\ref{fig:multidim-incoherent-compcost-EPIC-LAMBDA-I} shows a covariant view of the total computational cost density per station per voxel \textbf{(sum of gridding, FFT, squaring and temporal averaging components)} for the EPIC architecture taken two station parameters at a time while keeping the rest fixed at the nominal values of a \texttt{LAMBDA-I}, \textbf{\texttt{SKA-low}, or a \texttt{SKA-low-core}} station denoted by grey dashed lines. The variation in the cost per voxel is dominated by the spatial FFT cost which is relatively constant in all these parameters and displays the same trends observed in Figure~\ref{fig:1D-incoherent-compcost-EPIC-LAMBDA-I}, namely, nearly constant in $N_\textrm{eps}$, $D_\textrm{e}$ and $D_\textrm{s}$. Figure~\ref{fig:multidim-incoherent-relative-compcost-EPIC-LAMBDA-I} shows the ratio of the computational cost of the EPIC architecture relative to the voltage beamforming architecture in Figure~\ref{fig:multidim-incoherent-compcost-BF-LAMBDA-I}. It is clearly noted that the cost of the EPIC architecture is significantly less than that of voltage beamforming for $N_\textrm{eps} \gtrsim 100$, and continues to decrease with increasing $N_\textrm{eps}$, implying that for full field of view imaging, EPIC has a clear and significant computational advantage over voltage \textbf{DFT} beamforming as $N_\textrm{eps}$ and packing density of elements in a station increase.

@@

\textbf{The EPIC architecture will be relevant for large cosmological arrays like PUMA that are considering a FFT-based correlator architecture. Obtaining visibilities from an EPIC architecture will involve an additional step of an inverse-FFT or an inverse-DFT back to the aperture plane from the accumulated images (not included in Figure~\ref{fig:1D-incoherent-compcost-EPIC-LAMBDA-I}). However, the inverse-FFT or DFT operation to obtain visibilities needs to be performed only once every accumulation interval, and is thus not expected to add significantly to the total cost budget.}
